Verifying the \shaThree\ Implementation from \openssl\ with the Software Analysis Workbench

### 2022-01-10 paper/conclusion.tex
Add revisions from Brett

The proof uses a second \cryptol\ specification based on the actual \openssl\ C implementation of \shaThree.
This specification reorders arrays in the specification to match the C standard following row major order.
It also includes table lookups for constant values in the computation but it computes those tables to show the provenance of the constants.
That povenance is needed later for the proof of equivalence with the C code.
\saw\ proves the input to output relationship equivalent between the \fips\ specification and this revised specification for the 256-bit digest on any message of size $0$ to $1088$ bits.

The final \saw\ proof shows the \keccak\ function in the revised specification is equivalent to the \keccak\ function in the actual C implementation.
\keccak\ is the core of the \shaThree\ algorithm and the most complex part.
The proof relies of the use of overrides.
\saw\ proves the \cryptol\ specifications of the five inner \keccak functions equivalent to their C counterparts and replaces them in the C code with the \sawcore\ from the \cryptol.
With the overrides, \saw\ proves the equivalence for \keccak\ in \cryptol\ and C over any input.

>>>>

The proof uses a second \cryptol\ specification based on the actual \openssl\ C implementation of \shaThree.
This specification reorders arrays in the specification to match the C standard following row major order.
It also includes table lookups for constant values in the computation but it computes those tables to show the provenance of the constants.
That provenance is needed later for the proof of equivalence with the C code.
\saw\ proves the input to output relationship equivalent between the \fips\ specification and this revised specification for the 256-bit digest on any message of size $0$ to $1088$ bits.

The final \saw\ proof shows the \keccak\ function in the revised specification is equivalent to the \keccak\ function in the actual C implementation.
\keccak\ is the core of the \shaThree\ algorithm and the most complex part.
The proof relies of the use of overrides.
\saw\ proves the \cryptol\ specifications of the five inner \keccak\ functions equivalent to their C counterparts and replaces them in the C code with the \sawcore\ from the \cryptol.
With the overrides, \saw\ proves the equivalence for \keccak\ in \cryptol\ and C over any input.

### 2022-01-10 paper/introduction.tex
Move imp to figs and move figs to proof section
Restructure sections and file names
Add background on SAW and SHA-3
Add clarification on the openSSl spec
Add revisions from Brett

  The reliability of this correctness check is often weakened because of the difficulty of performing that visual review.  
  Many implementations of the SHA algorithms go through major rewrites and optimizations which make visual inspection of the code nearly impossible. (Not sure if that’s too strong a claim).
  This work bridges that gap between optimized c code, and the NIST standard.
  \input{proofFigure.tex}
\noindent \textbf{Contributions}:
This paper details the following contributions.
\begin{compactitem}
  \item  A Cryptol specification for \texttt{SHA3} that directly corresponds to the \texttt{FIPS} 202 specification and passes the test vectors provided by the same specification.
  \item A Cryptol specification based on the OpenSSL implementation of \texttt{SHA3} that uses a reordered state vector and precomputed tables in Keccak for performance gains and passes test vectors
  \item A \texttt{SAW} proof that the \texttt{FIPS} 202 specification is equivalent to the OpenSSL specification
  \item A \texttt{SAW} proof that the OpenSSL Keccak specification is equivalent to the OpenSSL C-implementation of the same function for all possible state vector input
  \item Early evidence that modular cryptographic specifications such as \texttt{SHA3} lend themselves to efficient \texttt{SAW} verification against C implementations as opposed to flattened specifications such as \texttt{SHA2} 
\end{compactitem}

\noindent \textbf{Limitations}
\begin{compactitem}
  \item Visual inspection and limited test vectors for (1) and (2)
  \item Only provide proof certificates for message sizes of X to Y for digest sizes of \textbf{FIXME} for (3)
  \item Visual inspection only for anything above Keccak for (3) in OpenSSL's C-code
\end{compactitem}

@@

Because of how NIST defines the standard, the keccak function is used for each digest size of the SHA3 algorithm, with the only significant differences being in the padding/sponging parts of the algorithm, which falls outside this work's scope.
This means that the proof of correctness for this function extends usefulness beyond just the SHA3-256 level which we focused on.
In our repository, we wrote the Cryptol specification to be able to run any of the SHA3 family of algorithms using the same keccak.
A simple extension of this work would be to take that already created and verified Cryptol specification, and prove it up against OpenSSL's slightly modified keccak versions.

>>>>

The reliability of this correctness check is often weakened because of the difficulty of performing that visual review.  
Many implementations of the SHA algorithms go through major rewrites and optimizations which make visual inspection of the code nearly impossible. (Not sure if that’s too strong a claim).
This work bridges that gap between optimized c code, and the NIST standard.

Give a brief history of \shaThree. Introduce briefly SAW and Cryptol.

% Contributions:

% (1) A Cryptol specification for SHA3 that passes test vectors and is readily comparable to the FIPS 202 specification
% (2) A Cpyptol specification for OpenSSL's implementation of SHA3 that passes test vectors and is readily comparable to the the OpenSSL C-code—this specification uses in the Keccak core pre-computed tables and a re-ordered state-vector
% (3) A SAW proof that the FIPS 202 specification is equivalent to the OpenSSL specification
% (4) A SAW proof that the OpenSSL Keccak specification is equivalent to the OpenSSL C-implementation of the same function for __**all possible state vectors**__

% Limitations:

% * Visual inspection and limited test vectors for (1) and (2)
% * Only provide proof certificates for message sizes of X to Y for digest sizes of … for (3)
% * Visual inspection only for anything above Keccak for (3) in OpenSSLs C-code

% @Ben: will you clarify your second point on the visual inspection for 2? I think what you are saying is that it is the FIPS 202 spec with the modified state vector that is what is used in OpenSSL, so there is no value in “visual inspection” in that sense. It which case I would restate (2) as

% (2) A Cpyptol specification based on OpenSSL implementation of SHA3 that uses a re-ordered state-vector and pre-computed tables in Keccak for performance gains that passes test vectors

% To a much lesser extent, I'd say another contribution is that we've shown that when a cryptographical spec is designed in a modular way (as the modern SHA3 is compared to SHA2) that facilitates formal verification---at least if using Cryptol and SAW. I believe we showed this by the fact that the SHA3 proofs complete successful for Yices, Z3, and ABC, whereas with our proofs for SHA2 only Yices completed.

\noindent \textbf{Contributions}:
This paper details the following contributions.
\begin{compactitem}
  \item  A Cryptol specification for \shaThree\ that directly corresponds to the \nist\ \fips\ specification and passes the test vectors provided by the same specification.
  \item A Cryptol specification based on the \openssl\ implementation of \shaThree\ that uses a reordered state vector and precomputed tables in Keccak for performance gains and passes test vectors
  \item A \saw\ proof that the \fips\ specification is equivalent to the OpenSSL specification
  \item A \saw\ proof that the \openssl\ \keccak specification is equivalent to the \openssl\ C-implementation of the same function for all possible state vector input
  \item Early evidence that modular cryptographic specifications such as \shaThree lend themselves to efficient \saw\ verification against C implementations as opposed to flattened specifications such as \shaTwo. 
\end{compactitem}

\begin{comment}
\noindent \textbf{Limitations}
\begin{compactitem}
  \item Visual inspection and limited test vectors for (1) and (2)
  \item Only provide proof certificates for message sizes of X to Y for digest sizes of \textbf{FIXME} for (3)
  \item Visual inspection only for anything above Keccak for (3) in OpenSSL's C-code
\end{compactitem}

@@

Because of how NIST defines the standard, the keccak function is used for each digest size of the SHA3 algorithm, with the only significant differences being in the padding/sponging parts of the algorithm, which falls outside this work's scope.
This means that the proof of correctness for this function extends usefulness beyond just the SHA3-256 level which we focused on.
In our repository, we wrote the Cryptol specification to be able to run any of the SHA3 family of algorithms using the same keccak.
A simple extension of this work would be to take that already created and verified Cryptol specification, and prove it up against OpenSSL's slightly modified keccak versions.
\end{comment}

### 2022-01-10 paper/openssl.tex
Add revisions from Brett

>>>>

The change to the state array structure and the added provenance giving the computation for lookup constants are the only changes to the \cryptol\ from the \fips.
These changes are sufficient for \saw\ to prove the input to output equivalence between the two \cryptol\ specifications for a fixed set of message sizes.
And they are sufficient to prove the equivalence to the actual C code in \openssl\ in the \keccak function for any arbitrary state.
The result adds more evidence to the ability to prove out implementations of cryptographic primitives when given specifications.

### 2022-01-10 paper/related.tex
Add revisions from Brett

Cryptol has been used to formally verify different implementations of the same cryptographic algorithm in the past.
As a demonstration of Cryptol's applications, Lewis et al \cite{crypt-hi} provided an portions of the Advanced Encryption Standard implemented in the language.
The language has also been used to verify equivalence between Skein Hash algorithm implementations \cite{hard-soft}.
Cryptol is not limited to hardware, as Browning and Weaver \cite{design-verif} have shown that the language can be tuned to rovide an abstraction of hardware.
They later provided a proof of equivalence between the FIPS AES specificiation and various optimized Cryptol implementations.
For certain tables, they initially demonstrated equivalence through mathematical principles.
These tables were also proven equivalent to less time-intensive statically defined tables used for the rest of their proofs.
As the OpenSSL implementation was highly optimized with similar static values, this paper mirrored the method Browning and Weaver outlined.
The language was also used to verify equivalence between the specification and hardware of the Verilog RTL's pseudorandom number generator \cite{pseudorandom}.

Decker et al's verification of OpenSSL's SHA2-256 implementation served as a direct predecessor to this work \cite{nfm-us}.
While SHA2 is also part of NIST's Secure Hash Algorithms, its internal structure differs vastly from the more complex and resistant SHA3.
The SHA2 proof, in turn, expanded the methodology by which Amazon's s2n HMAC was verified \cite{s2n-blog}.
While SHA3 is not implemented in s2n, the HMAC verificatiprovided a on utilized both a high-level and low-level cryptol specification in a similar manner to this work.
Both works compared equivalent C and Cryptol specification through SAW.

Other tools have been used on a wide variety of cryptographic algorithms.
STP \cite{stp}, a SAT-based procedure, has been used to verify implementation of AES, DES, SHA-1, and other block cipher functions \cite{auto}.
Isabelle \cite{isabelle} has aided in formal verification of networking protocols \cite{formal-protocols}.
The Coq \cite{coq} Proof assistant was used to prove equivalence between the OpenSSL, mbedTLS, and NIST implementations of HMAC-DRBG \cite{verified-correctness}.
Using Jasmin \cite{jasmin}, Barthe et al \cite{hi-assurance} showed that implementations of the ChaCha20 and Poly1305 were equivalent to their specifications.
Vale has been used to verify the SHA-256 and AES implementations \cite{vale}.
More close to home, the VeriHash framework demonstrated a bug in the RHash \cite{rhash} implementation of SHA3-256 \cite{hash-functions}.

>>>>

\cryptol\ has been used to formally verify different implementations of the same cryptographic algorithm in the past.
As a demonstration of \cryptol, Lewis \emph{et al} provided portions of the \emph{Advanced Encryption Standard} (AES) implemented in the language \cite{crypt-hi}.
The language has also been used to verify equivalence between Skein Hash algorithm implementations \cite{hard-soft}.
\cryptol\ is not limited to hardware, as Browning and Weaver \cite{design-verif} have shown that the language can be tuned to provide an abstraction of hardware.
They later provided a proof of equivalence between the FIPS AES specification and various optimized \cryptol\ implementations.
For certain tables, they initially demonstrated equivalence through mathematical principles.
These tables were also proven equivalent to less time-intensive statically defined tables used for the rest of their proofs.
As the OpenSSL implementation was highly optimized with similar static values, this paper mirrored the method Browning and Weaver outlined.
The language was also used to verify equivalence between the specification and hardware of the Verilog RTL's pseudorandom number generator \cite{pseudorandom}.

Decker et al's verification of OpenSSL's \shaTwo-256 implementation served as a direct predecessor to this work \cite{nfm-us}.
While \shaTwo is also part of the \nist\ Secure Hash Algorithms, its internal structure differs vastly from the more complex and resistant \shaThree.
The \shaTwo proof, in turn, expanded the methodology by which Amazon's s2n HMAC was verified \cite{chudov,s2n-blog}.
While \shaThree is not implemented in s2n, the HMAC verification provided utilized both a high-level and low-level \cryptol\ specification in a similar manner to this work.
Both works compared equivalent C and \cryptol\ specification through \saw.
Decker's work is inspired by the \openssl verification of HMAC \cite{beringer}.

Other tools have been used on a wide variety of cryptographic algorithms.
STP \cite{stp}, a SAT-based procedure, has been used to verify implementation of AES, DES, \shaOne, and other block cipher functions \cite{auto}.
Isabelle \cite{isabelle} has aided in formal verification of networking protocols \cite{formal-protocols}.
The Coq \cite{coq} Proof assistant was used to prove equivalence between the OpenSSL, mbedTLS, and NIST implementations of HMAC-DRBG \cite{verified-correctness}.
Using Jasmin \cite{jasmin}, Barthe et al \cite{hi-assurance} showed that implementations of the ChaCha20 and Poly1305 were equivalent to their specifications.
Vale has been used to verify the \shaTwo-256 and AES implementations \cite{vale}.
More close to home, the VeriHash framework demonstrated a bug in the RHash \cite{rhash} implementation of SHA3-256 \cite{hash-functions}.
